### 2018-08-27 slowtests/slowtests-boston.Rmd
update tests

---
title: "Tests: Bouston Housing Data"
author: "Brandon M. Greenwell"
output: html_document
---

```{r setup}
# Load required packages
library(ggplot2)
library(grid)
library(pdp)

# Test function
boston_test <- function(object, ...) {
  trn <- subset(boston, select = -cmedv)
  pd1 <- partial(object, pred.var = "lstat", train = trn, ...)
  pd2 <- partial(object, pred.var = c("lstat", "rm"), chull = TRUE,
                 train = trn, ...)
  pd3 <- partial(object, pred.var = "rm", ice = TRUE, train = trn, ...)
  pd4 <- partial(object, pred.var = "rm", ice = TRUE, center = TRUE,
                 train = trn, ...)
  grid.arrange(
    autoplot(pd1, rug = TRUE, train = trn, main = "PDP for lstat"),
    autoplot(pd2, main = "PDP for lstat and rm"),
    autoplot(pd3, main = "ICE curves for rm", alpha = 0.1),
    autoplot(pd4, main = "c-ICE curves for rm", alpha = 0.1),
    ncol = 2, top = textGrob(deparse(substitute(object)),
                             gp = gpar(fontsize = 20, font = 3))
  )
}
```

# Cubist

## Package: `Cubist`

```{r Cubist_cubist}
boston_cubist <- Cubist::cubist(x = subset(boston, select = -cmedv),
                                y = boston$cmedv, committees = 100)
boston_test(boston_cubist)
```

# Decision trees

## Package: `rpart`

```{r rpart_rpart}
boston_rpart <- rpart::rpart(cmedv ~ ., data = boston)
boston_test(boston_rpart)
```

## Package: `party`

```{r party_ctree}
boston_ctree <- party::ctree(cmedv ~ ., data = boston)
boston_test(boston_ctree)
```

## Package: `partykit`

```{r partykit_ctree}
boston_ctree2 <- partykit::ctree(cmedv ~ ., data = boston)
boston_test(boston_ctree2)
```

# Bagging

## Package: `ipred`

```{r ipred_bagging}
set.seed(101)
boston_ipred <- ipred::bagging(cmedv ~ ., data = boston, nbagg = 500)
boston_test(boston_ipred)
```

# Random forests

## Package: `randomForest`

```{r randomForest_randomForest}
set.seed(101)
boston_rf <- randomForest::randomForest(cmedv ~ ., data = boston)
boston_test(boston_rf)
```

## Package: `party`

```{r party_cforest}
set.seed(101)
boston_crf <- party::cforest(cmedv ~ ., data = boston)
boston_test(boston_crf, quantiles = TRUE)
```

## Package: `ranger`

```{r ranger_ranger}
set.seed(101)
boston_ranger <- ranger::ranger(cmedv ~ ., data = boston)
boston_test(boston_ranger)
```

# Boosting

## Package: `gbm`

```{r gbm_gbm}
set.seed(101)
boston_gbm <- gbm::gbm(cmedv ~ ., data = boston, distribution = "gaussian",
                       n.trees = 5000, interaction.depth = 3, shrinkage = 0.001,
                       cv.folds = 5)
best.iter <- gbm::gbm.perf(boston_gbm, method = "OOB", plot.it = FALSE)
boston_test(boston_gbm, n.trees = best.iter)
```

## Package: `xgboost`

```{r xgboost_xgboost}
set.seed(101)
boston_xgb <- xgboost::xgboost(
  data = data.matrix(subset(boston, select = -cmedv)), label = boston$cmedv,
  objective = "reg:linear", nrounds = 5000, max_depth = 3, eta = 0.001,
  colsample_bytree = 0.8, subsample = 0.5,
  save_period = NULL, verbose = 0
)
boston_test(boston_xgb)
```

# Neural networks

## Package: `nnet`

```{r nnet_nnet}
set.seed(101)
boston_nnet <- nnet::nnet(cmedv ~ ., data = boston, size = 6, decay = 0.1,
                          linout = TRUE, trace = FALSE)
boston_test(boston_nnet)
```

# Support vector machines

## Package: `e1071`

```{r e1071_svm}
boston_svm <- e1071::svm(cmedv ~ ., data = boston, type = "eps-regression")
boston_test(boston_svm)
```

## Package: `kernlab`

```{r kernlab_ksvm}
boston_ksvm <- kernlab::ksvm(cmedv ~ ., data = boston, type= "eps-svr")
boston_test(boston_ksvm)
```

# Linear/generalized linear models

## Package: `stats`

```{r stats_lm}
boston_lm <- lm(cmedv ~ . ^ 2, data = boston)
boston_test(boston_lm)
```

```{r stats_glm}
boston_glm <- glm(cmedv ~ . ^ 2, data = boston)
boston_test(boston_glm)
```

# Multivariate adaptive regression splines

## Package: `earth`

```{r earth_earth}
boston_earth <- earth::earth(cmedv ~ ., degree = 3, data = boston)
boston_test(boston_earth)
```

## Package: `mda`

```{r mda_mars}
boston_mars <- mda::mars(x = subset(boston, select = - cmedv), 
                         y = boston$cmedv, degree = 3)
boston_test(boston_mars)
```